---
title: "sporo3perc_analysis"
author: "T Attenborough"
date: "21/03/2022"
output: html_document
---

Load packages
```{r}
library(Seurat) #Seurat v4.0.2
library(dplyr)
library(patchwork)
library(here)
library(ggplot2)
library(clustree)
library(ggrepel)
library(leiden)
#sessionInfo()
```

Import the lit marker document 
```{r}
lit_markers <- read.csv(here("/Users/ta13/Documents/my_lit_markers.csv"))
lit_markers$Gene_ID <- gsub('\\_', '-', lit_markers$Gene_ID) #replace dashes with underscores
```

Now, import the final sporocyst data object from Carmen,and 
```{r}
sporocysts <- readRDS(file = "/Users/ta13/data/Seurat_Object/Final_objects/sporo_dataset_final.RDS")
```

Now, import the final somule data object from Carmen
```{r}
somules <- readRDS(file = "/Users/ta13/data/Seurat_Object/Final_objects/somules.RDS")
```

Filter out cells with more than 3% mito gene expression
```{r}
sporocysts3 <- subset(sporocysts, subset = percent.mt < 3)
dim(sporocysts3) #shows number of genes and number of cells
```

This leaves us with 601 cells.

Re-cluster with only 3percent mito
```{r}
sporocysts3 <- FindVariableFeatures(sporocysts3, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(sporocysts3)
sporocysts3 <- ScaleData(sporocysts3, features = all.genes)

sporocysts3 <- RunPCA(sporocysts3, features = VariableFeatures(object = sporocysts3))

# Examine and visualize PCA results a few different ways
print(sporocysts3[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(sporocysts3, dims = 1:2, reduction = "pca")

DimPlot(sporocysts3, reduction = "pca")

DimHeatmap(sporocysts3, dims = 1, cells = 500, balanced = TRUE)

sporocysts3 <- JackStraw(sporocysts3, num.replicate = 100, dims =50)
sporocysts3 <- ScoreJackStraw(sporocysts3, dims = 1:50)
JackStrawPlot(sporocysts3, dims = 1:50)
ElbowPlot(sporocysts3)
```

```{r}
#Check how many PCs to use for the data using Mathews Molecular cross validation
source("code.R")
mcv = molecularCrossValidation(sporocysts3@assays$RNA@counts, varGenes = VariableFeatures(sporocysts3),normalisation = minimalSeuratV3)

plotMCV(mcv)
```

This indicates the use of 12 PCs

PLot new cluster, and some QC data
```{r}
sporocysts3 <- FindNeighbors(sporocysts3, dims = 1:12)
sporocysts3 <- FindClusters(sporocysts3, resolution = 0.4,  algorithm = 1)
sporocysts3 <- RunUMAP(sporocysts3, dims = 1:12)
DimPlot(sporocysts3, reduction = "umap", label = TRUE)+ NoLegend()
ggsave("sporocysts3_pc12_res0.4.pdf", width = 15, height = 15, units = c('cm'))


FeatureScatter(object = sporocysts, feature1 = "nFeature_RNA", feature2 = "percent.mt", group.by="ident")

FeatureScatter(object = sporocysts3, feature1 = "nFeature_RNA", feature2 = "percent.mt", group.by="ident")
table(sporocysts3@meta.data$RNA_snn_res.0.4)
```
Try with res of 3 to capture neurons
```{r}
sporocysts3 <- FindNeighbors(sporocysts3, dims = 1:12)
sporocysts3 <- FindClusters(sporocysts3, resolution = 3)
sporocysts3 <- RunUMAP(sporocysts3, dims = 1:12)
DimPlot(sporocysts3, reduction = "umap", label = TRUE)+ NoLegend()
ggsave("sporocysts3_pc12_res3.pdf", width = 15, height = 15, units = c('cm'))


FeatureScatter(object = sporocysts, feature1 = "nFeature_RNA", feature2 = "percent.mt", group.by="ident")

FeatureScatter(object = sporocysts3, feature1 = "nFeature_RNA", feature2 = "percent.mt", group.by="ident")
table(sporocysts3@meta.data$RNA_snn_res.3)
```

Use clustree to find optimal cluster res
```{r}
# Select a range of resolutions
resolution.range <- seq(from = 0, to = 3, by = 0.5)

# Find clusters using a range of resolutions
sporocysts3 <- Seurat::FindClusters(object = sporocysts3, resolution = resolution.range)

clustree(sporocysts3)
ggsave("sporocysts3.clust_12PC_0_to_3_tree_SCT.pdf", width = 15, height = 25, units = c('cm'))
```

Use somule data to initially project annotations onto miracidia run1 sorted data
```{r}
DefaultAssay(sporocysts3) <- "RNA"

#This is for label transfer
pair.anchors <- FindTransferAnchors(reference = somules, query = sporocysts3, dims = 1:25, features = intersect(VariableFeatures(somules), VariableFeatures(sporocysts3)))

predictions <- TransferData(anchorset = pair.anchors, refdata = Idents(somules), dims = 1:25)

sporocysts3 <- AddMetaData(sporocysts3, metadata = predictions)
DimPlot(sporocysts3, group.by = 'predicted.id', label = T, label.size = 2)
ggsave("sporocysts3_labels_from_somules.pdf")
```


Cluster 12 gives the neuron cluster

Print the cell IDs of the cluster 12 cells
```{r}
#Idents(object = sporocysts3)
WhichCells(object = sporocysts3, ident = '12')
neuronIDs <- WhichCells(object = sporocysts3, ident = '12')
```

Re-run the clustering at 0.4 res
```{r}
sporocysts3 <- FindNeighbors(sporocysts3, dims = 1:12)
sporocysts3 <- FindClusters(sporocysts3, resolution = 0.4,  algorithm = 1)
sporocysts3 <- RunUMAP(sporocysts3, dims = 1:12)
DimPlot(sporocysts3, reduction = "umap", label = TRUE)+ NoLegend()
ggsave("sporocysts3_pc12_res0.4.pdf", width = 15, height = 15, units = c('cm'))
```

How many clusters?
```{r}
table(Idents(sporocysts3))
```

Rename the neuron cluster as 5

```{r}
Idents(sporocysts3, cells = c(neuronIDs)) <- '5'
head(Idents(sporocysts3))
table(Idents(sporocysts3))
```

PLot the UMAP
```{r}
DimPlot(sporocysts3, reduction = "umap", label = TRUE)+ NoLegend()
```

```{r}
sporocysts3 <- RenameIdents(object = sporocysts3, 
                               "0" = "Muscle",
                               "1" = "Tegument 1",
                               "2" = "Stem/germinal",
                               "3" = "Parenchymal",
                               "4" = "Tegument 2",
                               "5" = "Neuron")

sporocysts3@active.ident <- factor(sporocysts3@active.ident, 
                            levels=c("Muscle",
                                     "Neuron",
                                     "Parenchymal",
                                     "Stem/germinal",
                                     "Tegument 1",
                                     "Tegument 2"))

DimPlot(sporocysts3, reduction = "umap", label = TRUE)+ NoLegend()
```

```{r}
sporocysts3[["may.ident"]] <- Idents(object = sporocysts3)
```


Add custom colour palette
```{r}
new_pal <- c("#ffb703","#007aaa","#107E7D","#c40c18","#fb8500","#7851a9")
plot1 <- DimPlot(sporocysts3, reduction = "umap", label = FALSE, repel = TRUE, label.box = FALSE) + NoLegend() +scale_color_manual(values = new_pal)
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75)
ggsave("fancyumapJun22.pdf", width = 15, height = 12, units = c('cm'))
```

Find all markers
```{r}
#find all markers
sporocysts3.markers_roctest <- FindAllMarkers(sporocysts3, only.pos = TRUE, min.pct = 0.0, logfc.threshold = 0.25, test.use = "roc", return.thresh = 0)

sporocysts3.top5_roc=sporocysts3.markers_roctest %>% group_by(cluster) %>% top_n(n = 5, wt = myAUC)# groups the markers by cluster, then picks the top 5 most differentiually expressed markers
write.csv(x=sporocysts3.markers_roctest, file="top5_markerz_roc_sporocysts3_jul5.csv")
```

Plot top 5 markers of each cluster by AUC
```{r}
sporocysts3.markers_roctest$gene <- gsub('\\_', '-', sporocysts3.markers_roctest$gene) #replace dashes in geneid with underscores
dotplot_features=sporocysts3.markers_roctest %>% group_by(cluster) %>% top_n(n = 5, wt = myAUC)

DotPlot(object = sporocysts3, features = unique(dotplot_features$gene)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_color_viridis(discrete = FALSE)
ggsave("sporocysts3_top5genes_myAUC_jul5.pdf", width = 25, height = 15, units = c('cm'))

dotplot_features=sporocysts3.markers_roctest %>% group_by(cluster) %>% top_n(n = 3, wt = myAUC)

DotPlot(object = sporocysts3, features = unique(dotplot_features$gene)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_color_viridis(discrete = FALSE)
ggsave("sporocysts3_top3genes_myAUC_jul5.pdf", width = 20, height = 15, units = c('cm'))

```

```{r}
carm_clust <- read.csv("/Users/ta13/R_tutorials/clean_miracidia/somule_seurat markers.csv", stringsAsFactors = FALSE, header = TRUE) 
carm_clust_min <- carm_clust %>% select (c(Gene.identifier, cluster, Description))

sporocysts3.markers_roctest$gene <- gsub('\\-', '_', sporocysts3.markers_roctest$gene) #replace dashes in geneid with underscores

sporocysts3.markers_roc_annotate <- sporocysts3.markers_roctest %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier")) 
write.csv(x=sporocysts3.markers_roc_annotate, file="sporocysts3.markers_roc_annotatejul6.csv") #save this as a csv
```

```{r}
dotplot_features=sporocysts3.markers_roc_annotate %>% group_by(cluster.x) %>% top_n(n = 5, wt = myAUC)
write.csv(x=dotplot_features, file="sporocysts3.markers_roc_annotatejul_top5.csv") #save this as a csv
```

Here I am going to find a list of marker genes for cluster 0. Carmen suggested genes which are expressed in more than 70% of cells in the cluster, and less than 30% of cells in other clusters.
```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Muscle")
#leaves 2768 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 28 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 6 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 6 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="0_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}
carm_clust <- read.csv("/Users/ta13/R_tutorials/clean_miracidia/somule_seurat markers.csv", stringsAsFactors = FALSE, header = TRUE)
carm_clust_min <- carm_clust %>% select (c(Gene.identifier, cluster, Description))

clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier")) 
write.csv(x=clustX_annotate, file="0_Markers_annotateMar22.csv") #save this as a csv

```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="viridis")
ggsave("0_Markers_annotateMar22_myAUC.pdf", width = 15, height = 12, units = c('cm'))
```

```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Tegument 1")
#leaves 666 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 78 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 49 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 49 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="1_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}
clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier")) 
write.csv(x=clustX_annotate, file="1_Markers_annotateMar22.csv") #save this as a csv

```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="cividis")
ggsave("1_Markers_annotateMar22_myAUC.pdf", width = 30, height = 12, units = c('cm'))
```

Here I am going to find a list of marker genes for cluster 2. 

```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Stem/germinal")
#leaves 1271 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 205 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 12 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 4 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="2_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}

clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier"))
write.csv(x=clustX_annotate, file="2_Markers_annotateMar22.csv") #save this as a csv

```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="cividis")
ggsave("2_Markers_annotateMar22_myAUC.pdf", width = 15, height = 12, units = c('cm'))
```

Here I am going to find a list of marker genes for cluster 3.
```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Parenchymal")
#leaves 961 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 27 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 8 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 8 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="3_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}
clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier"))
write.csv(x=clustX_annotate, file="3_Markers_annotateMar22.csv") #save this as a csv
```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="cividis")
ggsave("3_Markers_annotateMar22_myAUC.pdf", width = 15, height = 12, units = c('cm'))
```

Here I am going to find a list of marker genes for cluster 4.

```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Tegument 2")
#leaves 660 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 102 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 9 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 9 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="4_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}
clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier"))
write.csv(x=clustX_annotate, file="4_Markers_annotateMar22.csv") #save this as a csv

```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="cividis")
ggsave("4_Markers_annotateMar22_myAUC.pdf", width = 15, height = 12, units = c('cm'))
```

Here I am going to find a list of marker genes for cluster 5.

```{r}
#Select just cluster 0
clustX <-filter(sporocysts3.markers_roctest, cluster == "Neuron")
#leaves 2502 genes

#Select just those in at least 70% of the clust0 cells
clustX <-filter(clustX, pct.1 >= 0.7)
#leaves 22 genes

#Select just those in 30% or less of the other cells
clustX <-filter(clustX, pct.2 <= 0.3)
#leaves 6 genes

#Select just those with at least myAUC of 0.7
clustX <-filter(clustX, myAUC >= 0.7)
#leaves 6 genes

clustX$gene <- gsub('\\-', '_', clustX$gene) #replace dashes in geneid with underscores

write.csv(x=clustX, file="5_Markers_Mar22.csv")

```

Attach some annotation info from somule markers
```{r}
clustX_annotate <- clustX %>% left_join(carm_clust_min, by = c("gene" = "Gene.identifier"))
write.csv(x=clustX_annotate, file="5_Markers_annotateMar22.csv") #save this as a csv

```

Make a dotplot of these genes
```{r}
features <- gsub('\\_', '-', clustX_annotate$gene) #replace dashes with underscores

DotPlot(object = sporocysts3, features = unique(features)) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+ scale_color_viridis(discrete = FALSE, option="cividis")
ggsave("5_Markers_annotateMar22_myAUC.pdf", width = 15, height = 12, units = c('cm'))
```

Check for some known marker genes
```{r}
#Intramammalian stem cell
features <- subset(lit_markers, Stage == "Intramammalian" & Population == "cpx+", select = Gene_ID)

# Feature plot - visualize feature expression in low-dimensional space
p1 = FeaturePlot(sporocysts3, features = features$Gene_ID) 
p1

ggsave("intramammalian_cpx+_3.pdf", width = 20, height = 30, units = c('cm'))


#there is no Smp_050220 in the stem cell clust
#this is intramallalian neural preogenitor markers

```

intramolluscan kappa markers
```{r}
features <- subset(lit_markers, Stage == "Intramolluscan" & Population == "kappa", select = Gene_ID)

# Feature plot - visualize feature expression in low-dimensional space
FeaturePlot(sporocysts3, features = features$Gene_ID)

ggsave("sporocysts_intramolluscan_kappa3.pdf", width = 35, height = 25, units = c('cm'))
```

```{r}
features <- subset(lit_markers, Stage == "Intramolluscan" & Population == "delta", select = Gene_ID)

# Feature plot - visualize feature expression in low-dimensional space
FeaturePlot(sporocysts3, features = features$Gene_ID)

ggsave("sporocysts_intramolluscan_delta3.pdf", width = 35, height = 25, units = c('cm'))
```

```{r}
features <- subset(lit_markers, Stage == "Intramolluscan" & Population == "phi", select = Gene_ID)

# Feature plot - visualize feature expression in low-dimensional space
FeaturePlot(sporocysts3, features = features$Gene_ID)

ggsave("sporocysts_intramolluscan_phi3.pdf", width = 35, height = 25, units = c('cm'))
```

Extract a list of all genes for comparison
```{r}
allGenesInSporos3 <- rownames(x = sporocysts3)

head(allGenesInSporos3)

allGenesInSporos3 <- data.frame(allGenesInSporos3)
allGenesInSporos3$allGenesInSporos3 <- gsub('\\-', '_', allGenesInSporos3$allGenesInSporos3) #replace dashes with underscores
write.csv(x=allGenesInSporos3, file="allGenesInSporos3.csv") #save this as a csv

```

Export the CSV of raw counts - just the stem cell cluster
```{r}

table(sporocysts3@meta.data$may.ident)
sporocysts3_subset <- subset(sporocysts3, idents = c("Stem/germinal"))
table(sporocysts3_subset@meta.data$may.ident)
counts <- sporocysts3_subset@assays$RNA@counts
write.csv(x = counts, file = '/Users/ta13/sporocysts3.stem.counts_2.csv')
write.csv(x = sporocysts3_subset@meta.data, file = '/Users/ta13/meta_fromsporocysts3.stem.csv')

write.csv(x = sporocysts3@meta.data, file = '/Users/ta13/meta_fromsporocysts3.csv')

```

Save that annotation information as metadata
```{r}
sporocysts3[["new.ident"]] <- Idents(object = sporocysts3)
```

```{r}
# load tidyverse packages
library(tidyverse)
# create cluster ID table
tibble(cellID = colnames(sporocysts3), clusterID = Idents(sporocysts3)) %>%
  # write out to a file, with today's date
  write_csv(file = sprintf("sporocysts3_cluster_mappings_%s.csv", Sys.Date()))

run1.combined_52_meta <- tibble::rownames_to_column(sporocysts3@meta.data, "cellID") %>%
  # write out to a file, with today's date
  write_csv(file = sprintf("sporocysts3_meta_%s.csv", Sys.Date()))
```


```{r}
saveRDS(sporocysts3, file = "sporocysts3.rds") #saves this version of the dataset

```
