---
title: "teg1_v_teg2_sporo"
author: "T Attenborough"
date: "2023-10-11"
output: html_document
---

Load packages
```{r}
library(Seurat) #Seurat v4.0.2
library(dplyr)
library(patchwork)
library(here)
library(ggplot2)
library(clustree)
library(ggrepel)
library(tidyverse)
library(topGO)
library(forcats)

#sessionInfo()
```

This saves today's date for later
```{r}
st=format(Sys.time(), "%Y-%m-%d") 
st
```

Now, import the threepercent sporocyst object
```{r}
sporocysts3 <- readRDS(file = "/Users/ta13/R_tutorials/sporocysts/sporocysts3.rds")
```

```{r}
table(sporocysts3@active.ident)
```

```{r}
sporocysts3.markers_roc_annotate <- read.csv("/Users/ta13/R_tutorials/sporocysts/sporocysts3.markers_roc_annotate_2022-07-18.csv", stringsAsFactors = FALSE, header = TRUE) #import the marker data from somule papaer
```

Import annotation information
```{r}
v9_genelist <- read.csv("/Users/ta13/R_git/r_teaching/WBPS17_genelist.csv", stringsAsFactors = FALSE, header = TRUE) #import the marker data from somule papaer
v9_genelist$X <- NULL 
```

```{r}
stem.response <- FindMarkers(sporocysts3, ident.1 = "Tegument 1", ident.2 = "Tegument 2", verbose = TRUE, min.pct=0.0, logfc.threshold=0.0)
head(stem.response, n = 15)
stem.response <- tibble::rownames_to_column(stem.response, "gene")
stem.response$gene <- gsub('\\-', '_', stem.response$gene) #replace dashes in geneid with underscores
stem.response <- stem.response %>% left_join(v9_genelist, by = c("gene" = "Name")) #check the top 5 marker genes are also in the somule paper
write.csv(x=stem.response, file="sporocysts3_teg1Vteg2.csv") #save this as a csv
```

```{r}
plot1 <- FeaturePlot(sporocysts3, features = "Smp-328760")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-328760-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))
```

I have split this file, so one has all the DEGs from delta/phi, and one has all the degs from kappa

Import the stem 1 DEGs
```{r}
teg1_deg <- read.csv("/Users/ta13/R_tutorials/sporocysts/sporocysts3_upin_teg1.csv", stringsAsFactors = FALSE, header = TRUE) 
```

Import the stem 1 DEGs
```{r}
teg2_deg <- read.csv("/Users/ta13/R_tutorials/sporocysts/sporocysts3_upin_teg2.csv", stringsAsFactors = FALSE, header = TRUE) 
```

SUBSET TO ONLY ADJ P VAL LESS THAN 0.001
```{r}
teg1_deg <- subset(teg1_deg, 
       p_val_adj < 0.001,       # rows
       select=c("gene", "p_val_adj")) # columns

teg1_deg$p_val_adj <- NULL

write.table(teg1_deg, file=paste0('teg1_deg_Markers_justgenes_', st, '.txt'), col.names = F, row.names = F,sep="\t", quote = FALSE)
```

```{r}
teg2_deg <- subset(teg2_deg, 
       p_val_adj < 0.001,       # rows
       select=c("gene", "p_val_adj")) # columns

teg2_deg$p_val_adj <- NULL

write.table(teg2_deg, file=paste0('teg2_deg_Markers_justgenes_', st, '.txt'), col.names = F, row.names = F,sep="\t", quote = FALSE)
```

```{r}
geneID2GO <- readMappings(file = "/Users/ta13/R_git/topgo_tutorial/test_GOterm_dict.txt", sep = " ") 
```

```{r}
geneUniverse <- names(geneID2GO) 
```


Extract a list of all genes for comparison
```{r}
DefaultAssay(sporocysts3) <- "RNA"

allGenesInSporo3 <- rownames(x = sporocysts3)

head(allGenesInSporo3)

allGenesInSporo3 <- data.frame(allGenesInSporo3)
allGenesInSporo3$allGenesInSporo3 <- gsub('\\-', '_', allGenesInSporo3$allGenesInSporo3) #replace dashes with underscores
write.csv(x=allGenesInSporo3, file="allGenesInSporo3.csv") #save this as a csv
```

Import the list of genes in the mira dataset
```{r}
listOfSporo3Genes <- read.csv("/Users/ta13/R_tutorials/sporocysts/allGenesInSporo3.csv", header=TRUE, stringsAsFactors=FALSE)
listOfSporo3Genes$X <- NULL

listOfSporo3Genes <- listOfSporo3Genes$allGenesInSporo3
```

Find the common elements from the full gene list from schisto, and the genes detected in sporo data
```{r}
sporo3_geneUniverse <- intersect(geneUniverse,listOfSporo3Genes)
```

```{r}
Rscript TopGO_mira.R teg1_deg_Markers_justgenes_2023-10-11.txt BP
Rscript TopGO_mira.R teg1_deg_Markers_justgenes_2023-10-11.txt MF
```

```{r}
Rscript TopGO_mira.R teg2_deg_Markers_justgenes_2023-10-11.txt BP
Rscript TopGO_mira.R teg2_deg_Markers_justgenes_2023-10-11.txt MF
```

Import the summary file
```{r}
topGO_summary_teg <- read.csv("/Users/ta13/R_tutorials/sporocysts/topGO_summary_sporos_teg_DE_withCC.csv", stringsAsFactors = FALSE, header = TRUE) #import the marker data from somule papaer
```

```{r}
str(topGO_summary_teg)

topGO_summary_teg$neglog10FDR
unique(topGO_summary_teg$Colour)
new_pal <- c("#fb8500","#7851a9")

#fct_reorder(topGO_summary_sporos$Term, topGO_summary_mira$Cell.type, min)
ggplot(data = topGO_summary_teg, aes(x = fct_reorder2(Term, neglog10FDR, Cell.type, min), y = neglog10FDR, fill = Cell.type)) + geom_bar(stat="identity") + 
  scale_colour_identity()+
       coord_flip()+theme_bw()+scale_fill_manual(values = new_pal)
ggsave("topGO_summary_teg_filtv9_plot1.pdf", width = 20, height = 35, units = c('cm'))

#fct_reorder(topGO_summary_sporos$Term, topGO_summary_sporos$Cell.type, min)
ggplot(data = topGO_summary_teg, aes(x = fct_reorder2(Term, neglog10FDR, Cell.type), y = neglog10FDR, fill = Cell.type)) + geom_bar(stat="identity") + 
  scale_colour_identity()+
       coord_flip()+theme_bw()+scale_fill_manual(values = new_pal)+ labs(fill = " ", x = " ", y = "-log10FDR")+theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=12), axis.title.x = element_text(size = 18),legend.key.size = unit(0.8, 'cm'), legend.text = element_text(size=10))
ggsave("topGO_summary_teg_filtv9_plot2.pdf", width = 20, height = 35, units = c('cm'))
```

```{r}
#fct_reorder(topGO_summary_sporos$Term, topGO_summary_sporos$Cell.type, min)
#fct_reorder(topGO_summary_sporos$Term, topGO_summary_sporos$Cell.type, min)
ggplot(data = topGO_summary_teg, aes(x = Term, y = neglog10FDR, fill = Cell.type)) + geom_bar(stat="identity") + 
  scale_colour_identity()+
       coord_flip()+theme_bw()+scale_fill_manual(values = new_pal)+ labs(fill = " ", x = " ", y = "-log10FDR")+theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=12), axis.title.x = element_text(size = 18),legend.key.size = unit(0.8, 'cm'), legend.text = element_text(size=10))
ggsave("topGO_summary_teg_filtv9_plot2.pdf", width = 15, height = 20, units = c('cm'))
```

```{r}
plot1 <- FeaturePlot(sporocysts3, features = "Smp-328760")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-328760-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))

plot1 <- FeaturePlot(sporocysts3, features = "Smp-121950")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-121950-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))

plot1 <- FeaturePlot(sporocysts3, features = "Smp-173350")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-173350-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))

plot1 <- FeaturePlot(sporocysts3, features = "Smp-348500")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-348500-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))

plot1 <- FeaturePlot(sporocysts3, features = "Smp-153370")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-153370-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))

plot1 <- FeaturePlot(sporocysts3, features = "Smp-121430")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-121430-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))
```


```{r}
plot1 <- FeaturePlot(sporocysts3, features = "Smp-090230")
LabelClusters(plot1, id = "ident", color = 'black', size =4, repel = T,  box.padding = 1.75, max.overlaps = Inf)
ggsave(paste("sporocysts3-teg-DE-Smp-090230-labels-",st, ".pdf", sep = ""), width = 25, height = 15, units = c('cm'))
```


